---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# ALTREP rep.int

The goal of this package is to learn about ALTREP and to prototype an ALTREP version of `rep.int` for scalar values. Not meant to be used in real situations.

Base's `rep.int(42, 10)` allocates an integer vector of 10 times 42.
Using ALTREP we can construct a version of `rep.int` (called `rep_int` :)) that
is lazy and does not allocate a full vector, yet still offers the same functionality. In case a fully expanded version of the vector is needed, it materializes it on the fly.

For example: the ALTREP system has a method to ask a data structure for a value
at a given index. Well, in this particular case the value is always the same, so we can answer this question in constant time (see below).

This proof of concept is based on the examples of [Luke Tierney](https://github.com/ALTREP-examples/Rpkg-mutable/blob/master/src/mutable.c) and the blog post
of [Romain François](https://purrple.cat/blog/2018/10/21/lazy-abs-altrep-cplusplus/).

## Installation

``` r
remotes::install_github("dirkschumacher/repint")
```

## Some examples

```{r example}
library(repint)
stopifnot(all.equal(rep.int(10, 1e6), rep_int(10, 1e6)))
bench::mark(
  rep.int(10, 1e6),
  rep_int(10, 1e6)
)
```

`head` and `sort` both work with ALTREP, so they can take advantage of the `rep_int` implementation.

```{r}
stopifnot(all.equal(head(sort(rep.int(10, 1e6))), head(sort(rep_int(10, 1e6)))))
bench::mark(
  head(sort(rep.int(10, 1e6))),
  head(sort(rep_int(10, 1e6)))
)
```

We also know that the sequence is sorted:

```{r}
is.unsorted(rep_int(42, 10))
```

It also supports materialization:

```{r}
a <- rep_int(42, 10)
.Internal(inspect(a))
```

The second line is the description of the materialized content. Currently Nil. If you want to learn more about that approach, take a look at Romain's [blog post]((https://purrple.cat/blog/2018/10/21/lazy-abs-altrep-cplusplus/).

Still lazy:

```{r}
b <- sort(a)
.Internal(inspect(a))
```

Still lazy:
```{r}
a[5]
.Internal(inspect(a))
```

Still lazy:
```{r}
a[4:6]
.Internal(inspect(a))
```


Materialized as soon as it is printed:

```{r}
b
.Internal(inspect(a))
```


